---
title: "BrailleR time analysis"
author: "James Thompson - 21011195"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BrailleR)
library(tidyverse)
library(microbenchmark)
```


To test all the different plots they will be run through a function like this which will return a sighted and blind number

```{r}
runTests <- function(plot, dataframe) {
  ggplot =  class(plot)[2] == "ggplot"
  sightedResult = blindAndSightedTime(plot, FALSE, ggplot)
  dataframe[nrow(dataframe)+1,] <- c(FALSE, ggplot, sightedResult)
  blindResult = blindAndSightedTime(plot, TRUE, ggplot)
  dataframe[nrow(dataframe)+1, ] <- c(TRUE, ggplot, blindResult)

  return(dataframe)
}

blindAndSightedTime = function(plot, withBrailleR, ggplot=FALSE) {
  if (!withBrailleR) {
    GoSighted()
    result = summary(microbenchmark(capture.output(print(plot))))
  } else {
    GoBlind()
    result = summary(microbenchmark(capture.output(print(plot))))
  }
  if (ggplot) {
    result$expr = class(plot$layers[[1]]$geom)[1]
  } else {
    result$expr = class(plot)[2]
  }
  
  return(result)
}
```

We will save the results in a dataframe
```{r}
# Results data frame
columns <- c("withBrailleR","ggplot", "expr", "min", "lq", "mean", "median", "uq", "max", "neval")
results <- data.frame(matrix(nrow = 0, ncol = length(columns))) 
colnames(results) <- columns
rm(columns)
```

# Base plots

```{r}
d = mtcars

boxplot = boxplot(d)

dotplot = dotplot(d$mpg)

hist = hist(d$mpg)

results = runTests(boxplot, results)
results = runTests(dotplot, results)
results = runTests(hist, results)

```

# ggplot

```{r fig.show='hide', warning = FALSE, messages = FALSE}

N <- 1e4
x <- rnorm(N)
randomData <- data.frame(x = x, y = rnorm(N, x))
rm(x,N)

# Hline
hline <- ggplot(data=randomData, aes(x=x, y=y)) + 
  geom_hline(yintercept = mean(randomData$y))
results = runTests(hline, results)

# point
point = ggplot(data=randomData, aes(x, y)) + 
  geom_point()
results = runTests(point, results)

# bar
data("mpg")
bar = ggplot(mpg, aes(class)) +
  geom_bar()
results = runTests(bar, results)

#line
data("economics")
line = ggplot(economics, aes(date, unemploy)) +
  geom_line()
results = runTests(line, results)

# boxplot
boxplot = ggplot(NULL, aes(x=randomData$x)) +
  geom_boxplot()
results = runTests(boxplot, results)

# smooth
smooth = ggplot(randomData, aes(x,y)) +
  geom_smooth()
results = runTests(smooth, results)

# ribbon
ribbon = ggplot(randomData, aes(x,y)) +
  geom_ribbon(aes(ymin = mean(randomData$y)-10, ymax = mean(randomData$y)+10))
results = runTests(ribbon, results)
```

Now that we have all of the data we can then run some analysis on it.

```{r}
avgRuntime = results |>
  group_by(withBrailleR) |>
  summarise(executionTime = mean(median))
avgRuntime$executionTime[2]/avgRuntime$executionTime[1]
```

We can see from this that it looks like BrailleR does make a difference.
It seems to slow down on average `r avgRuntime$executionTime[2]/avgRuntime$executionTime[1]`.

However it is better to compare the increase between each plot type, then average across the lot. This is because the differences can be an order of magnitude different.
```{r}
avgIncrease = results |>
  group_by(withBrailleR) |>
  mutate(increase = median/lag(median)) |>
  drop_na() |>
  ungroup() |>
  group_by(ggplot) |>
  summarise(increase = mean(increase))
avgIncrease
```
We can see something funny here where base plots actually run quicker when using BrailleR then without.

The average increase seems to be about 50% for ggplot.

But where does that extra time come from
```{r}
GoBlind()
profvis::profvis({print(hline)}, interval=0.005)
Sys.sleep(2)
GoSighted()
profvis::profvis({print(hline)}, interval=0.005)
GoBlind()
profvis::profvis({print(smooth)}, interval=0.005)
```

We can see that it pretty clean cut that the increase in time comes form VI.ggplot call. This then runs through.
This means that the results calculated before have some merit to them.