# Mischeanolous working

This will ahve various bug fixes and issues resovles that dont fit into any of the other sections

## GITHUB nightmare

I started working quite quickly.

Didnt really know the procedure properly and so forked repor and worked like usual.

This meant that I was branching and merging into my master as opposed to the orginal repo.

Had to reset my master back to original and slowly copy changes from bracnhes into clean PRs

Once that was done then I could carry on with clean work on expand limits and beyond.

## Wrong title names

29/11/22 The problem here is the one with issue #24.

It is that the base plot hist isnt picking up on the main title properly.

```{r}
x <- hist(airquality$Ozone, main = "")
y <- hist(airquality$Ozone)
VI(x)
VI(y)
```

I think that this problem is right across the board for the base plots. As it looks like they are grabbing x$ExtraArgs$main rather than x\$main. They should check first for a given name then just go to the default name.

### Solution 1

A potential solution could be to have a function that gabs the name of a graph regardless of the type. As I believe the base plots are all similar in that way.

```{r}
getGraphName = function(graph) {
  if (length(graph$main) > 0 && nchar(gsub(" ", "",graph$main, fixed=T)) != 0) {
    paste("With the title:",graph$main)
  } else if (nchar(graph$ExtraArgs$main) > 0) {
    paste("With the title:",graph$ExtraArgs$main)
  } else {
    "With no title"
  }
}
```

Will this work with all the types of supported base plots

```{r}
x = rnorm(1e4)

boxplotWith = boxplot(x, main="This is the title")
getGraphName(boxplotWith)

boxplotWithout = boxplot(x)
getGraphName(boxplotWithout)

boxplotNone = boxplot(x, main="   ")
getGraphName(boxplotNone)


dotplotWith = with(airquality, dotplot(Ozone~Month,main="This is the title"))
getGraphName(dotplotWith)

dotplotWithout = with(airquality, dotplot(Ozone~Month))
getGraphName(dotplotWithout)

dotplotNone = with(airquality, dotplot(Ozone~Month, main=" "))
getGraphName(dotplotNone)


histWith = hist(x, main="This is the cool title")
getGraphName(histWith)

histWithout = hist(x)
getGraphName(histWithout)

histNone = hist(x, main="  ")
getGraphName(histNone)
```

Im going to say that this works for the three base plots and I will implement it into the package.

This is probably a good point to note that I think it could be a good idea to switch them all over to mustache templates.

Here is testing using that method integrated into the package. I am using the plots from above

```{r}
VI(boxplotWith)
VI(boxplotWithout)
VI(boxplotNone)

VI(dotplotWith)
VI(dotplotWithout)
VI(dotplotNone)

VI(histWith)
VI(histWithout)
VI(histNone)
```

Besides more github troubles with accidentally pushing too early this has all gone smoothly

## issue 62 timeseries not calling ggplot

6/12/22 We have the problem taht in the TimeSeriesPlot function there simply isnt a call to the ggplot where there should be.

We can add it in but we will need more variables and currently it is a bit messy with the naming of the variables.

```{r}
attach(airquality)
op = par(mfcol=c(3,2))
plot(as.ts(Wind), ylab="Wind", col=4)
test1 = TimeSeriesPlot(x=Wind, col=4, base=TRUE)
test1 #does the plot method work?
plot(as.ts(Ozone), ylab="Ozone")
test2 = TimeSeriesPlot(x=Ozone, base=TRUE)
test2 # does the plot method work?
par(op)
detach(airquality)

#ggplot tests
#Normal data
basic = TimeSeriesPlot(airquality, x=airquality$Temp)
basic
#ts data
withTitle = TimeSeriesPlot(UKDriverDeaths,  main = "UK driver deaths", sub = "For the month Januaray")
withTitle
withLabs = TimeSeriesPlot(sunspots, ylab="Number of spots", xlab="Year")
withLabs

xdata = rnorm(153)
test = TimeSeriesPlot(airquality, x=xdata)
test
```

### Problem

There was a n error that the examples in the help files that I added didnt work properly. It couldnt find the data which was being sent.

## issue 63

6/12/22

Ths is working with the help pages of both VI and Describe.

I have split them out. Haven't added much text as there was already alot. I have made the Authors more representative of who has done work.

## issue #67 build failure

It seems that there is a problem when importing the XML package on the github build server.

I will reduce the imports and see if that helps.

I did this and it did help.

But then the pkgdown started failing.

I have fixed it up with a suggestion from the issue on he github action repo.

## Switching to Bookdown

5/01/2022

At the start of this new year I am trying to organise my notes.

I thought it would be best to put them into a bookdown as it will let me separate it out a little bit.

### Scope issues

There have been some referencing issues where I am not putting all the information for it to run in the rmd file.

This in particular with using internal functions so I have used ::: to make some of these methods accessible.

### SVG problems

There were more problems that I was facing with the baseR integration with SVGThis and AddXML. As we are considering that baseR graphics are deprecated when it comes to BrailleR I have simply not evaluated these in the rendering of this book.

### Timeseries analysis

I ahve a largish rmd file which is all of the timeseries analysis. This takes a very long time to render so becuase of that I removed it form the book. It is something that I have in my files and can render if needed.

### Resolution

All in all it went well and it is now sorted.

## Adding in pre-commits

I have added some pre-commits to this repo. It will means that whenever a file is committed that code will ahve to pass some checks as well as be formatted.

This should help prevent little types and such breaking the code. It will also start to enforce a style guide on the code so it is indented as you would expect.
